---
title: "MA 677 Take-Home Assignment"
author: "Hsueh-Pin Liu"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=F, include=F}
library(dplyr)
```

# Problem 1

Consider three boxes:

A box containing two gold coins.

A box containing two silver coins.

A box containing one gold coin and a silver coin.

You choose a box at random and withdraw one coin at random. It's a gold coin. What is the probability that the other coin in the same box is also a gold coin?

## Simulation

Besides the basic calculation method that I did it correct in class, the answer should be $\frac{2}{3}$, and I decide to use simulate the experiment for 100000 times and see the results.

```{r}
set.seed(1)
n <- 100000
x <- c("1G","1G","2S","2S","3G","3S")
y <- sample(x,size=n,replace=T)
y <- as.data.frame(y)
m <- count(y,y)
m
#To satisfy the problem, we should calculate the ratio of "1G" to "1G"+"3G".
print(33244/(33244+16766))
```

As a conclusion the answer of this problem is approximately equal to $\frac{2}{3}$ by using simulation. And then I want to explore something about statistical paradoxes, after searching a lot of information, I find an interesting one.

## Birthday Paradox

Imagine that there are 23 students in a class, what's the probability that two of them have the same birthday? Some people might think that there are 365 days in a year, so the probability is $\frac{23}{365}$ which is even smaller than 10%. But in fact, the correct formula should be $1-\frac{364}{365}*\frac{363}{365}*...*\frac{343}{365}$, and it is calculated as follows:

```{r}
print(1-prod(343:364)/365^22)
```

Surprisingly, we got a result that is over 0.5, which means that the probability two of them having the same birthday is higher than the probability that all of them have a different birthday!

# Problem 3

Suppose that a random variable Y has a probability density function given by $f(y)=ky^3e^{-(y/2)}$ for $y>0$.

Find the value of k and calculate $E[y^2]$.

## Solution 1

Before I think about the gamma function, I use integration by parts to solve the problem. $\int u dv=uv-\int vdu$, so the function can be calculated as follows.

```{=tex}
\begin{center}

$\int_{0}^{\infty}f(y)dy=\int_{0}^{\infty}ky^3e^{-\frac{y}{2}}dy=k(y^3(-2e^{-\frac{y}{2}})+\int_{0}^{\infty}6y^2e^{-\frac{y}{2}}dy$

$\int_{0}^{\infty}6y^2e^{-\frac{y}{2}}dy=6y^2(-2e^{-\frac{y}{2}})+\int_{0}^{\infty}24ye^{-\frac{y}{2}}dy$

$\int_{0}^{\infty}24ye^{-\frac{y}{2}}dy=24y(-2e^{-\frac{y}{2}})+\int_{0}^{\infty}48e^{-\frac{y}{2}}dy$

\end{center}
```

So we can get the final result.

```{=tex}
\begin{center}

$\int_{0}^{\infty}f(y)dy=k(y^3+6y^2+24y+48)(-2e^{-\frac{y}{2}})|_{0}^{\infty}=96k=1$

$k=\frac{1}{96}$

\end{center}
```

And we can calculate $E[y^2]$ using the same method.

```{=tex}
\begin{center}

$E[y^2]=\int_{0}^{\infty}f(y)*y^2dy=\frac{1}{96}(y^5+10y^4+80y^3+480y^2+1920y+3840)(-2e^{-\frac{y}{2}})|_{0}^{\infty}=\frac{7680}{96}=80$

$E[y^2]=80$

\end{center}
```

This is calculation without using the gamma function, so it's more complex and takes more time.

## Solution 2

Let's consider $\Gamma(4,2)$.


```{=tex}
\begin{center}

$f(y)=\frac{y^3e^{-\frac{y}{2}}}{2^4\Gamma(4)}=\frac{y^3e^{-\frac{y}{2}}}{96}$
\end{center}
```

And we know that 

```{=tex}
\begin{center}

$\int_{0}^{\infty}\frac{y^3e^{-\frac{y}{2}}}{96}=1$

\end{center}
```

So

```{=tex}
\begin{center}

$k=\frac{1}{96}$

\end{center}
```

In order to calculate $E[Y^2]$, we can consider $\Gamma(6,2)$

```{=tex}
\begin{center}

$f(y)=\frac{y^5e^{-\frac{y}{2}}}{2^6\Gamma(6)}=\frac{y^5e^{-\frac{y}{2}}}{7680}$

$\int_{0}^{\infty}\frac{y^5e^{-\frac{y}{2}}}{7680}=1$

$E[y^2]=\int_{0}^{\infty}f(y)*y^2dy=\int_{0}^{\infty}\frac{y^5e^{-\frac{y}{2}}}{96}dy=80\int_{0}^{\infty}\frac{y^5e^{-\frac{y}{2}}}{7680}dy=80$

\end{center}
```






# Problem 5

Suppose that the random variables X and Y are such that $E[X] = 4$, $E[Y] = −1$, $V[X] = 2$,
and $V[Y] = 8$.

(a) What is the largest possible value of $Cov(X,Y)$?

(b) Assuming $\rho_{X,Y}$= 0.3, find $V(X,Z)$ where $Z = X+bY$ and $b ∈ R$.

## Solution

(a) Because 

```{=tex}
\begin{center}

$\rho_{X,Y}=\frac{Cov(X,Y)}{\sqrt{V(X)V(Y)}},\rho_{X,Y}\leq 1$

\end{center}
```

So we can have

```{=tex}
\begin{center}

$Cov(X,Y)\leq\sqrt{V(X)V(Y)}=\sqrt{2*8}=4$

$\max Cov(X,Y)=4$

\end{center}
```

(b)

```{=tex}
\begin{center}

$V(X,Z)=V(X,X+bY)=V(X,X)+bV(X,Y)=V(X,X)+0.3b\sqrt{V(X)V(Y)}=2+1.2b$

\end{center}
```

## Deep Thinking

From question(a), I have two problems.

1) When can $Cov(X,Y)$ get the largest value?

2) If $\rho_{X,Y}=0$, does it mean that there's not any relation between them?

### 1)

When $Cov(X,Y)$ get the largest value, we can know that $\rho_{X,Y}=1$, which also means that there are constants $a,b$ such that $P\{Y=aX+b\}=1$, and to say this in an easy way, there is a linear relationship between $Y$ and $X$.

### 2)

In fact, we can consider the following example. Assume that $X_1=\sin{x}$, $X_2=\cos{x}$,we know that $\rho_{X1,X2}=0$, but we can't say there are not any relation between them since $X_1^2+X_2^2=1$, so when $\rho_{X,Y}=0$ we can only get the conclusion that they don't have a linear relationship.

# Problem 6

Suppose that the number of minutes you have to wait for a bus is uniformly distributed on the interval $[0, 15]$. You wait for the bus 3 times. Find the probability that your longest wait is less than 10 minutes.

## Simulation

Besides the basic calculation method that I did it correct in class, the answer should be $\frac{8}{27}$, and I decide to use simulate the experiment for 100000 times and see the results.

```{r}
n <- 100000
x <- matrix(runif(n*3,min=0,max=15),nrow=n,ncol=3)
max <- apply(x,1,max)
print(sum(max<10)/n)
```

As a conclusion the answer of this problem is approximately equal to $\frac{8}{27}$ by using simulation. And for the situation mentioned in the problem, I have a similar example. When shopping in a market, if you are buying some products that you haven't buy before, A has 7 left on the shelf and B has only 2 left, which one will you choose? Probably most of us will choose B because we think it's more popular. We can actually use this idea to sell products with high inventory.








